### 2023-02-08 article.tex
Add response rates

The code used in experiments is just that, 'data about the data collection 
process', as it commonly downloads data, selects relevant subsets in those
downloaded data and performs statistical tests to generate results.
Code, hence, is paradata and this chapter explores and illustrates the
consequence of that premise.

@@

The reasons why useful paradata matters,
are data reuse, research reproducibility, 
contextual understanding of disciplinary differences,
Open Science and insights into 
(scientific) knowledge production \cite{huvila2022improving}.
This chapter discusses these reasons below.

@@

This paper uses genetic epidemiology as a specific example,
to illustrate that the code is paradata,
that this type of paradata is relevant, as it helps 
to assess the correctness of the results,
where the correctness of the results is relevant for 
science in general and healthcare in particular.
However, any field that uses computation in its experiments
could be used as an example.

@@

Going back to the example scientists in the introduction,
where one does and the other does not publish the code,
it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
only 192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where still only 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert \cite{haibe2020importance}.

@@

Genetic epidemiology is a field within biology that, among 
others, measures the spread of heritable traits,
as well as how these come to be.
For example, we know that lactose intolerance in adults is
caused by a decline in the production of lactose-degrading enzymes,
and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
The trait is caused by the genetic make-up, 
or genotype 
(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
of a person.
The trait, also called phenotype (again, see Table \ref{tab:definitions}), 
in this example is lactose intolerance at adult age,
yet any human property, such as weight or height can be studied.
Genetic epidemiology studies the spread of genotypes and phenotypes,
as well as the relation between those two.
When an association between genotype and phenotype is found,
these associations are subsequently used to 
create gene panels, where the gene causing 
an association is measured specifically, to, among others,
detect people at risk for the associated phenotype.

The rest of this section describes a genetic epidemiology study 
in more detail, with special focus on the computational experiment.

@@

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).

>>>>

The code used in computational experiments is just that, 
'data about the data collection 
process', as it commonly downloads data, selects relevant subsets in those
downloaded data and performs statistical tests to generate results.
Code, hence, is paradata and this chapter explores and illustrates the
consequence of that premise.

@@

There are multiple reasons why useful paradata matter. 
Most obvious is that having useful paradata gives an understanding how
data is produced.
This insight helps researchers from different fields to understand each other
and collaborate.
Additionally, useful open data is needed for Open Science to convincingly show
its benefits.
Finally, in computational fields, 
it can help understand how scholarly knowledge is produced
\cite{huvila2022improving}.

@@

This paper uses genetic epidemiology as a specific example,
to illustrate that the code is paradata,
and that this type of paradata is relevant.
However, any field that uses computation in its experiments
could be used as an example.

@@

Going back to the example scientists in the introduction,
where one does and the other does not publish the code,
it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
only 192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where still only 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

Genetic epidemiology is a field within biology that
tries to determine the mechanism and spread of heritable traits.
For example, we know that lactose intolerance in adults is
caused by a decline in the production of lactose-degrading enzymes,
and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
The trait is caused by the genetic make-up, or 'genotype'
(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
of a person.
The trait, also called 'phenotype' (again, see Table \ref{tab:definitions}), 
in this example is lactose intolerance at adult age,
yet any human property, such as weight or height can be studied.
When an association between genotype and phenotype is found,
these associations are subsequently used to 
create so-called 'gene panels', 
where the gene causing 
an association is measured specifically, to, among others,
detect people at risk for the associated phenotype.
The rest of this section describes a genetic epidemiology study 
in more detail, with special focus on the computational experiment.

@@

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To be more precise, it was code that read the data,
subsetted the data, removed outliers, 
performed the statistics and generated the plots.
All these steps are described in the paper, 
but the code itself is unpublished
and the paper does not mention any way to obtain it.
The reponse rate of correnspondence authors with a request for data is around 50\% 
\cite{manca2018non, teunis2015corresponding} 
(note that for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
